The Gender Pay Gap in the General Social Survey

### 2023-08-31 Paper/paper.Rmd
proofread papper

---
title: "The Gender Pay Gap in the General Social Survey"
author: "Alex Amaguaya, Lea Bergmann"
cols_authors: 3
subtitle: "Statistical Learning"
deadline: "tomorrow"
type: "Term Paper"
date: "today"
supervisors: "Prof. Dr. Thomas Deckers, Karolina Gliszczynska"
output:
  pdf_document:
    keep_tex: yes
    template: template.tex
    fig_caption: yes
    citation_package: biblatex
    number_sections: true
toc: true
lot: true
lof: true
graphics: true
biblio-title: References
fontsize: 11pt
geometry: lmargin=2.5cm,rmargin=2.5cm,tmargin=2.5cm,bmargin=2.5cm
biblio-files: ref.bib
classoption: a4paper
---

@@

The prediction of individuals' wages has been an issue in previous literature. As wages are relevant to companies they want to know how predict wages [@Chakraborti]. Some papers make use of classification techniques to predict wages [@Chakraborti]. As we have numeric values of real wages and not a categorical variable we do not make use of classification techniques but prefer linear regression techniques such as OLS and Lasso regressions as well as random forests. We prefer these techniques as they handle categorical variables and outliers well, which are features of our data [@Cutler]. Wages were already predicted using random forests by [@Eichinger]. They find that using random forests results in better predictions than using linear models. Additionally, they show that gender barely effects the wage predictions. Likewise, the boosting methods used by [@Chakrabarty2018] in salary level classification problems showed better performance compared to a traditional algorithm such as SVM. Furthermore, [@BonaccoltoToepfer1612] suggests the use of a penalization approach for high dimensional problems in his research about Gender wage inequality.

@@

Before modeling we conduct a data preparation process. The initial data set has 54,842 observations and the outcome variable `realrinc` has 38.55% of missing values. The imputation of the target variable is discarded to avoid generating a bias in the modeling and therefore we decided to eliminate these records from the original data set. After this, the data set contains 33,702 observations. Furthermore, it is impossible to deduce the occupation for some observations (`occ10 = 9997`) and we remove these from the data set. Finally, the data set has 33,244 observations and we use it for the modeling process.

This data set still had some missing values for the regressors, e.g., number of children, age, marital status, degree level, and others. So, we use the `mice` package with five variables for the imputation process. The features *age, occupational prestige and number of children* are imputed using the predictive mean matching (pmm) method, and we use the polytomous logistic regression method for the factor variables (*education and marital status*). Due to we use a value of 6 in the `m` parameter of the `mice` function, the imputation process generates 6 different data sets. This parameter refers to the number of imputed data sets. After this, we apply the calculation of the mean and the majority vote to aggregate the data sets into one. Mean estimation is applied for numerical variables and majority voting for categorical variables

@@

After the imputation process, we created some interaction variables. First, we estimated some interaction features using only the numerical variables. Then, we transformed the numerical variables (*age, child and prestg*) into categorical variables by separating them into ranges. The ranges created of these variables were: *age between 18 and 30, age between 31 and 50, and age greater than 50, child between 0 and 2, child between 3 and 5, child greater than 5; and prestg10 between 16 and 30, prestg10 between 31 and 50, prestg10 greater than 50*. After this, we transform these new categorical variables and the initial categorical variables (*marital status, education, etc.*) into dummy variables. Finally, the interaction variables are estimated with the group of dummy variables mentioned above (the interaction consists of the multiplication between 2 dummy variables). The inclusion of the interaction of the numeric variables are validate using a linear model, and the results show that the p-values of the parameters for the interactions are less than 0.05.

@@

Furthermore, we estimate a correlation matrix using Spearman's rank correlation coefficient with some numeric variables in order to support the previous results. The correlation results are exposed in the figure below, and show that *prestg10*, *age* and *interaction between age and prest10* have the highest positive correlation with *realinc*, meanwhile the dummy variable *female* has a moderate negative correlation with the target variable.

@@

Due to the correlation assesses the linear relationship between two variables and it only uses numerical features, we decided to estimate the Predictive Power Score (PPS). It focuses on these limitations and allows us to have a predictive power of a regressor on a target variable. This measure is a normalized metric and its values are between 0 (no predictive power) and 1 (perfect predictive power). The PPS is calculated by comparing the performance of a model that uses a regressor `x` to predict target variable `y` versus the performance of a base model (naive). When the target variable is a numerical feature it is common to use a Decision Tree Regressor to predict `y` based on `x`, while in categorical cases a Decision Tree Classifier is used. On the other hand, base models are often defined as mean prediction or random or modal prediction for regression and classification problems, respectively. The PPS results are presented in the figure below, and show that *prestg10*, *wrkstat*, *educcat* and *occrecode* have the highest predictive power over *realinc* relative to the other features. These results help and provide some insights in the data exploration process, but we must beware of the limitations of this approach.

@@

After the data preparation and feature engineering processes, we have a final data set with 33,244 observations and 421 features. Due to the large number of interaction variables, we decided to reduce them and use only the interaction between a group of variables (*occrecode*, *educcat* and categorical variable of *prestg10*). After this, we develop the modeling process with 100 variables (numerical and interaction variables). The following figure shows the overview of the group of variables that were estimated and used in the models.

@@

As the dataset contains many variables (361) and even our selection of dummy variables for the right-hand side of the regression leaves us with 53 variables, we make use of regression techniques for subset selection. We use lasso, ridge and elastic net regressions. These methods allow us to fit a model that contains all variables. Then the model regularizes the coefficient estimates and shrinks them towards zero. In effect, we reduce the estimates' variance which improves the model fit.

@@

Hence, the coefficient estimates in a ridge regression minimize the sum of the residual sum of squares (RSS) and the sum of squared coefficients multiplied with a tuning parameter $\lambda$. We estimate $\lambda$ using cross-validation.

@@

According to the literature review, tree-based models perform considerably well in predicting income ranges (classification problems) [@Chakrabarty2018]. For this reason, we use the random forest algorithm for modeling. In this experiment, income without logarithm transformation, initial variables and some interaction variables are used.
Due to the amount of hyperparameters in random forest, we use the `lightgbm` and `mlr3` packages to model and tune the hyper parameter in a time-optimal way. After many attempts, we found the best hyper parameters that produced the lowest error. The hyperparameters tuned were `max_depth`, `num_leaves`, `num_iterations`, etc, and the search space of them is showed below. In addition, we used a random search approach in the tuning process with a cross-validation of 3 folds.

@@

Once we found the model along with the hyperparameters with lowest error in the cross-validation, we proceeded to evaluate this model on the test dataset and obtained a RMSE of 31,745.54. The hyperparameters of the best model are shown in the following figure.

![Best hyperparameters for Random Forest](includes/RF_hyper.png)

In addition, we can analyze the features importance within the modeling process using random forest, and these results are exposed in the figure below.

@@

Unlike the linear models presented previously, random forest does not predict negative values when logarithmic transformation is not used. We carry out an experiment with a random forest but using the logarithmic transformation in the target variable and keeping the same configuration (CV of 3 folds and random search approach in the tuning process).

```r
learner.wlog = lrn("regr.lightgbm",
              boosting = "rf",
              objective = "regression",
              max_depth = to_tune(seq(7,12,1)),
              num_leaves = to_tune(seq(120,140,1)),
              max_bin = to_tune(seq(45,60,1)),
              num_iterations  = to_tune(seq(50,90,1)),
              min_data_in_leaf = to_tune(seq(20,35,1)), 
              min_data_in_bin = to_tune(seq(35,50, 1)),
              feature_fraction_bynode = to_tune(seq(0.4,0.7,0.1)),
              bagging_fraction = to_tune(seq(0.6,0.9,0.1)),
              bagging_freq = to_tune(seq(3,12,1)),
              feature_fraction = to_tune(seq(0.8,0.9,0.1)),
              convert_categorical = TRUE,
              force_col_wise = TRUE,
              verbose = 1,
              num_threads = 5,
              seed = 123
)

The best model using the logarithmic transformation had different hyperparameters relative to the previous experiment and obtained an RMSE of 34,293.33 on the test dataset. This result shows that the error of this model is greater than the best model without the transformation (31,745.54).

@@

Gradient Boosting also has a similar amount of hyperparameters as random forest and we decided to keep a similar configuration as the previous algorithm to find the best model with the hyperparameters (random search approach in the tuning process with 3-fold cross-validation). This first experiment with Gradient Boosting uses the target variable without logarithmic transformation.

@@

We tried many combinations of hyperparameters and found the best model that produced the least error. In Gradient Boosting we adjusted some hyperparameters such as `max_depth`, `num_leaves`, `min_data_in_leaf`, `feature_fraction` and others. Once the best model was found, it was evaluated on the test data set and an RMSE of 29,207.55 was obtained. The hyperparameters of the best model are shown in the following table.

![Best hyperparameters for Gradient Boosting](includes/GB_hyper.png)

Gradient Boosting also provides information on the importance of features during the modeling process in this experiment. The following figure shows the 20 most important variables according to the Gini metric.

@@

Furthermore, we carried out an additional experiment with gradient boosting using the logarithm transformation on the target variable.

```r
learner.gb.wlog = lrn("regr.lightgbm",
                 boosting = "gbdt",
                 objective = "regression",
                 max_depth = to_tune(seq(5, 12, 1)),
                 max_bin = to_tune(seq(35,60,1)), 
                 num_leaves = to_tune(seq(19,30,1)),
                 min_data_in_leaf = to_tune(seq(30,50,1)),
                 min_data_in_bin = to_tune(seq(20,45,1)),
                 feature_fraction = to_tune(seq(0.3,0.8,0.1)),
                 feature_fraction_bynode = to_tune(seq(0.7,0.9,0.1)),
                 learning_rate = to_tune(seq(0.01, 0.1, 0.01)), 
                 num_iterations  = to_tune(seq(60,80,1)), 
                 lambda_l1 = to_tune(seq(0.01, 0.1, 0.01)),
                 lambda_l2 = to_tune(seq(0.3, 0.45, 0.01)),
                 convert_categorical = TRUE,
                 force_col_wise = TRUE,
                 verbose = 1,
                 num_threads = 5,
                 seed = 123
)

@@

The best model using the logarithmic transformation had a RMSE of 34,385.47 on the test dataset. This result shows that the error of this model is greater relative to the best model without the transformation (29,207.55). These results show that in tree-based methods, Gradient Boosting has a dominant performance over Random Forest. In other words, the error of Gradient Boosting is about 8% smaller than the error of Random Forest when the logarithmic transformation is not used, but the errors of these models are larger than those of the linear models.

@@

Generalized additive model (GAM) could be considered the next step of the linear model because it allows combining nonlinear functional forms of the features along with the linear part. This approach is based on relating the nonlinear predictors with a link function $g(.)$ of the expected values $E(y)$, and the following formulas offer more details about this approach.

$$

y \sim \text{ExpoFam}(\mu, \text{etc})\\
E(y) = \mu \\
g(\mu) = b_0 + f(x_1) + f(x_2) + \dots + f(x_p)

$$

The main difference with Generalized Lineal Models (GLM) is that this approach allows linear features to aggregate smooth functions $f(.)$ of the required variables and thus establish nonlinear relationships between regressors and the dependent variable.
This models uses the `gam` and `mlr3` packages to model for the modeling. The adjusted hyperparameter was `gamma` and is responsible for producing a smoother term, and a additional penalization on the smooth term was aggregated with the parameter `select`. In addition, we used a 3-fold cross-validation in the training stage for this algorithm. This experiment uses the logarithmic transformation of the target variable, the interaction variables, smooth terms for the features *age* and *prestg10* with cubic regression splines `cr` and a value of 9 for the knots `k` (number of sections into which a feature is divided), and the smooth tensor product for the variables *age*, *prestg10* and *child* uses a thin-plate regression spline `tp` and a knot value of 1. The last smooth term is considered as the interaction between two features to predict the target variables.

```r
form.str.gam <- paste("log_realrinc ~ s(age, bs = 'cr', k = 9) + s(prestg10, bs = 'cr', k=9) + 
                        ti(age,prestg10) + ti(age,childs) +",
                       paste( x_cols_dummys, collapse = " + ") )
form.format.gam <- as.formula(form.str.gam)
learner.gam.wlog.smooth.tensor = lrn("regr.gam",
                                     family = 'gaussian',
                                     select = TRUE,
                                     gamma = to_tune(seq(1,3,0.5)),
                                     formula = form.format.gam
)
```

The best model obtained a RMSE of 25,856.76 on the test dataset. This result shows that the error of this model is slightly lower in relation to the linear model that also uses the logarithmic transformation (25,909.82). In other words, it is 0.20% better than the initial linear model. The figure below exposes the approximate significance of the smooth terms, and the results show that all the smooth terms used in the model have a p-value of less than 0.01.

![Approximate significance of smooth terms](includes/aprox_smooth.png)

@@

\begin{tabular}{l c} 
\textbf{Method} & \textbf{RMSE} \\ 
Linear Regression & 25,506.05 \\ 
Logistic Regression & 25,909.82 \\ 
Ridge & 26,062.58 \\ 
Lasso & 25,918.25 \\
Elastic Net & 25,918.81 \\ 
Random Forest & 31,745.54 \\ 
Gradient Boosting & 29,207.55 \\ 
Generalized Additive Model & 25,856.76 \\
\end{tabular}

\medskip
Conclusively, the linear regression yields the lowest RMSE and thus appears to be the best prediction method. Still, it delivers negative predictions of the wage. Therefore, we prefer the logistic regression as the best prediction method.

>>>>

---
title: "The Gender Pay Gap in the General Social Survey"
author: "Alex Amaguaya, Lea Bergmann"
cols_authors: 3
subtitle: "Statistical Learning"
deadline: "31.08.2023"
type: "Term Paper"
date: "31.08.2023"
supervisors: "Prof. Dr. Thomas Deckers, Karolina Gliszczynska"
output:
  pdf_document:
    keep_tex: yes
    template: template.tex
    fig_caption: yes
    citation_package: biblatex
    number_sections: true
toc: true
lot: true
lof: true
graphics: true
biblio-title: References
fontsize: 11pt
geometry: lmargin=2.5cm,rmargin=2.5cm,tmargin=2.5cm,bmargin=2.5cm
biblio-files: ref.bib
classoption: a4paper
---

@@

The prediction of individuals' wages has been an issue in previous literature. As wages are relevant to companies they want to know how to predict wages [@Chakraborti]. Some papers make use of classification techniques to predict wages [@Chakraborti]. As we have numeric values of real wages and not a categorical variable we do not make use of classification techniques but prefer linear regression techniques such as OLS and Lasso regressions as well as random forests. We prefer these techniques as they handle categorical variables and outliers well, which are features of our data [@Cutler]. Wages were already predicted using random forests by [@Eichinger]. They find that using random forests results in better predictions than using linear models. Additionally, they show that gender barely effects the wage predictions. Likewise, the boosting methods used by [@Chakrabarty2018] in salary level classification problems showed better performance compared to a traditional algorithm such as support vector machines (SVM). Furthermore, @BonaccoltoToepfer1612 suggest the use of a penalization approach for high dimensional problems in his research about gender wage inequality.

@@

Before modeling we conduct a data preparation process. The initial data set has 54,842 observations and the outcome variable `realrinc` has 38.55% of missing values. The imputation of the target variable is discarded to avoid generating a bias in the modeling and therefore we eliminate these records from the original data set. After this, the data set contains 33,702 observations. Furthermore, it is impossible to deduce the occupation for some observations (`occ10 = 9997`) and we remove these from the data set. Finally, the data set has 33,244 observations and we use it for the modeling process.

This data set still has some missing values for the regressors, e.g., number of children, age, marital status, degree level, and others. Thus, we use the `mice` package with five variables for the imputation process. The features *age, occupational prestige and number of children* are imputed using the predictive mean matching (pmm) method, and we use the polytomous logistic regression method for the factor variables (*education and marital status*). As we use a value of 6 in the `m` parameter of the `mice` function, the imputation process generates 6 different data sets. This parameter refers to the number of imputed data sets. After this, we apply the calculation of the mean and the majority vote to aggregate the data sets into one. Mean estimation is applied for numerical variables and majority voting for categorical variables

@@

After the imputation process, we create interaction variables. First, we estimate some interaction features using only the numerical variables. Then, we transform the numerical variables (`age, childs and prestg10`) into categorical variables by separating them into ranges. The ranges created of these variables were: `age between 18 and 30, age between 31 and 50`, `age greater than 50; childs between 0 and 2, childs between 3 and 5, childs more than 5; prestg10 between 16 and 30, prestg10 between 31 and 50, prestg10 greater than 50`. After this, we transform these new categorical variables and the initial categorical variables (`marital status`, `education`, etc.) into dummy variables. Finally, we estimate the interaction variables with the group of dummy variables mentioned above (the interaction consists of the multiplication between two dummy variables). The inclusion of the interactions of the numeric variables is validate using a linear model, and the results show that the p-values of the parameters for the interactions are below 0.05.

@@

Furthermore, we estimate a correlation matrix using Spearman's rank correlation coefficient with some numeric variables in order to support the previous results. The correlation results are exposed in figure 2, and show that `prestg10`, `age` and `interaction between age and prest10` have the highest positive correlation with `realrinc`, meanwhile the dummy variable `female` has a moderate negative correlation with the target variable.

@@

As the correlation assesses the linear relationship between two variables and it only uses numerical features, we additionally estimate the Predictive Power Score (PPS). It overcomes these limitations and allows us to assess the predictive power of a regressor on a target variable. This measure is a normalized metric and its values are between 0 (no predictive power) and 1 (perfect predictive power). The PPS is calculated by comparing the performance of a model that uses a regressor `x` to predict target variable `y` versus the performance of a base model (naive). When the target variable is a numerical feature it is common to use a Decision Tree Regressor to predict `y` based on `x`, while in categorical cases a Decision Tree Classifier is used. On the other hand, base models are often defined as mean prediction or random or modal prediction for regression and classification problems, respectively. The PPS results are presented in figure 3, and show that `prestg10`, `wrkstat`, `educcat` and `occrecode` have the highest predictive power of `realrinc` relative to the other features. These results provide some insights in the data exploration process, but we must beware of the limitations of this approach.

@@

After the data preparation and feature engineering processes, we have a final data set with 33,244 observations and 421 features. Due to the large number of interaction variables, we reduce them and use only the interaction between a group of variables (`occrecode`, `educcat` and categorical variable of `prestg10`). After this, we develop the modeling process with 100 variables (numerical and interaction variables). Figure 4 shows the overview of the group of estimated variables that are used in the models.

@@

As the data set contains many variables (361) and even our selection of dummy variables for the right-hand side of the regression leaves us with 53 variables, we make use of regression techniques for subset selection. We use lasso, ridge and elastic net regressions. These methods allow us to fit a model that contains all variables. Then the model regularizes the coefficient estimates and shrinks them towards zero. In effect, we reduce the estimates' variance which improves the model fit.

@@

Hence, the coefficient estimates in a ridge regression minimize the sum of the residual sum of squares (RSS) and the sum of squared coefficients multiplied with a tuning parameter $\lambda$. We estimate $\lambda$ using CV.

@@

According to the literature review, tree-based models perform considerably well in predicting income ranges (classification problems) [@Chakrabarty2018]. For this reason, we use the random forest algorithm for modeling. For this method we use income without logarithm transformation, initial variables and some interaction variables. Due to the amount of hyperparameters in random forest, we use the `lightgbm` and `mlr3` packages to model and tune the hyperparameter in a time-optimal way. We tune the hyperparameters `max_depth`, `num_leaves`, `num_iterations`, etc. to find the best hyperparameters that produce the lowest error. The search space of the hyperparameters is showed below. In addition, we use a random search approach in the tuning process with a CV of 3 folds.

@@

After identifying the model along with the hyperparameters with lowest error in CV, we proceed to evaluate this model on the test data set and obtain an RMSE of 31,745.54. The hyperparameters of the best model are shown in figure 5.

```{r, echo=FALSE, out.width="50%", fig.cap="Best Hyperparameters for Random Forest"}
knitr::include_graphics("includes/RF_hyper.png")
```

In addition, we analyze the features importance within the modeling process using random forest. These results are exposed in the figure 6.

@@

Unlike the linear models presented previously, random forest does not predict negative values when logarithmic transformation is not used. Still, we additionally use a random forest with the logarithmic transformation in the target variable and keep the same configuration (CV of 3 folds and random search approach in the tuning process).

``` r
learner.wlog = lrn("regr.lightgbm",
              boosting = "rf",
              objective = "regression",
              max_depth = to_tune(seq(7,12,1)),
              num_leaves = to_tune(seq(120,140,1)),
              max_bin = to_tune(seq(45,60,1)),
              num_iterations  = to_tune(seq(50,90,1)),
              min_data_in_leaf = to_tune(seq(20,35,1)), 
              min_data_in_bin = to_tune(seq(35,50, 1)),
              feature_fraction_bynode = to_tune(seq(0.4,0.7,0.1)),
              bagging_fraction = to_tune(seq(0.6,0.9,0.1)),
              bagging_freq = to_tune(seq(3,12,1)),
              feature_fraction = to_tune(seq(0.8,0.9,0.1)),
              convert_categorical = TRUE,
              force_col_wise = TRUE,
              verbose = 1,
              num_threads = 5,
              seed = 123
)
```

The best model using the logarithmic transformation has different hyperparameters relative to the previous experiment and obtains an RMSE of 34,293.33 on the test data set. Thus, the error of this model is greater than the best model without the transformation (31,745.54).

@@

Gradient Boosting has a similar amount of hyperparameters as random forest and we use a similar configuration as the previous algorithm to find the best model with the hyperparameters (random search approach in the tuning process with 3-fold CV). The first model with Gradient Boosting uses the target variable without logarithmic transformation.

@@

Our best model uses a combination of hyperparamters that produces the least error. We adjust some hyperparameters such as `max_depth`, `num_leaves`, `min_data_in_leaf`, `feature_fraction` and others. We evaluate the best model with the test data set and obtain an RMSE of 29,207.55. The hyperparameters of the best model are shown in figure 7.

```{r, echo=FALSE, out.width="50%", fig.cap="Best Hyperparameters for Gradient Boosting"}
knitr::include_graphics("includes/GB_hyper.png")
```

Gradient Boosting also provides information on the importance of features during the modeling process in this experiment. Figure 8 shows the 20 most important variables according to the Gini metric. `Gender` is after work status the most important variable, giving additional evidence for a gender pay gap.

@@

Furthermore, we estimate a gradient boosting model using the logarithm transformation on the target variable.

``` r
learner.gb.wlog = lrn("regr.lightgbm",
                 boosting = "gbdt",
                 objective = "regression",
                 max_depth = to_tune(seq(5, 12, 1)),
                 max_bin = to_tune(seq(35,60,1)), 
                 num_leaves = to_tune(seq(19,30,1)),
                 min_data_in_leaf = to_tune(seq(30,50,1)),
                 min_data_in_bin = to_tune(seq(20,45,1)),
                 feature_fraction = to_tune(seq(0.3,0.8,0.1)),
                 feature_fraction_bynode = to_tune(seq(0.7,0.9,0.1)),
                 learning_rate = to_tune(seq(0.01, 0.1, 0.01)), 
                 num_iterations  = to_tune(seq(60,80,1)), 
                 lambda_l1 = to_tune(seq(0.01, 0.1, 0.01)),
                 lambda_l2 = to_tune(seq(0.3, 0.45, 0.01)),
                 convert_categorical = TRUE,
                 force_col_wise = TRUE,
                 verbose = 1,
                 num_threads = 5,
                 seed = 123
)

@@

The best model using the logarithmic transformation has an RMSE of 34,385.47 on the test data set. Thus, the error of this model is greater relative to the best model without the transformation (29,207.55). These results show that in tree-based methods, Gradient Boosting has a dominant performance over Random Forest. In other words, the error of Gradient Boosting is about 8% smaller than the error of Random Forest when the logarithmic transformation is not used. Nevertheless, the errors of these models are larger than those of the linear models.

@@

Generalized additive models (GAM) could be considered the next step of the linear model because they allow combining nonlinear functional forms of the features along with a linear part. This approach is based on relating the nonlinear predictors with a link function $g(.)$ of the expected values $E(y)$. The following formulas offer more details about this approach.

$$
y \sim \text{ExpoFam}(\mu, \text{etc.})
$$ $$
E(y) = \mu
$$ $$
g(\mu) = b_0 + f(x_1) + f(x_2) + \dots + f(x_p)
$$ The main difference with Generalized Lineal Models (GLM) is that this approach allows linear features to aggregate smooth functions $f(.)$ of the required variables and thus establish nonlinear relationships between regressors and the dependent variable. We use the `gam` and `mlr3` packages to model the GLM. The adjusted hyperparameter is `gamma` and is responsible for producing a smoother term. An additional penalization on the smooth term is aggregated with the parameter `select`. In addition, we use a 3-fold CV in the training stage for this algorithm. We use the logarithmic transformation of the target variable, the interaction variables, smooth terms for the features `age` and `prestg10` with cubic regression splines `cr` and a value of 9 for the knots `k` (number of sections into which a feature is divided). The smooth tensor product for the variables `age`, `prestg10` and `child` uses a thin-plate regression spline `tp` and a knot value of 1. The last smooth term is considered as the interaction between two features to predict the target variables.

``` r
form.str.gam <- paste("log_realrinc ~ s(age, bs = 'cr', k = 9) 
                                    + s(prestg10, bs = 'cr', k=9) + 
                        ti(age,prestg10) + ti(age,childs) +",
                       paste( x_cols_dummys, collapse = " + ") )
form.format.gam <- as.formula(form.str.gam)
learner.gam.wlog.smooth.tensor = lrn("regr.gam",
                                     family = 'gaussian',
                                     select = TRUE,
                                     gamma = to_tune(seq(1,3,0.5)),
                                     formula = form.format.gam
)
```

The best model obtaines an RMSE of 25,856.76 on the test data set. Thus, the error of this model is slightly lower in relation to the linear model that also uses the logarithmic transformation (25,909.82). In other words, it is 0.20% better than the initial linear model. Figure 9 exposes the approximate significance of the smooth terms, and the results show that all the smooth terms used in the model have a p-value of less than 0.01.

```{r, echo=FALSE, out.width="80%", fig.cap="Approximate Significance of Smooth Terms"}
knitr::include_graphics("includes/aprox_smooth.png")
```

@@

```{=tex}
\begin{tabular}{l c} 
\textbf{Method} & \textbf{RMSE} \\ 
Linear Regression & 25,506.05 \\ 
Logistic Regression & 25,909.82 \\ 
Ridge & 26,062.58 \\ 
Lasso & 25,918.25 \\
Elastic Net & 25,918.81 \\ 
Random Forest & 31,745.54 \\ 
Gradient Boosting & 29,207.55 \\ 
Generalized Additive Model & 25,856.76 \\
\end{tabular}
```
\medskip

Conclusively, the linear regression yields the lowest RMSE and thus appears to be the best prediction method. Still, it delivers negative predictions of the wage. Therefore, we prefer the logistic regression. Comparing the RMSE of the logistic model to the one of the GAM, the GAM (which presents a combination of linear and nonlinear functional forms of the features) produces the lowest RMSE. Hence, the GAM presents the best prediction method for real wages (in our data set).

Moreover, the different prediction methods consider `gender` as an important determinant for predicting the individuals' real wage. Still, as we do not use any causal methods other unobserved factors might influence the wage more than `gender`.

\pagebreak

\addcontentsline{toc}{section}{References}
\printbibliography[title = References]
\cleardoublepage

\begin{refsection}
\nocite{R-base}
\nocite{R-Studio}

\end{refsection}
